Methods for Projecting Population Growth

### 2018-05-08 final/tex/Paper.tex
Finished :D

\begin{figure}[t!]
\includegraphics[width=80mm]{images/world-pop.png}
\includegraphics[width=80mm]{images/world-gdp.png}
\caption{World population and GDP modeled with parameters  $\alpha_1 = 0.3\%$, $\alpha_3 k_1 = \frac{5.2}{10^{6}}\frac{Hab.}{US \$}$, $\alpha_2 = \frac{-55}{10^{17} US\$}$, $\beta_1 = 3.6\%$, and $\beta_2 = \frac{-2}{10^{22} Hab.}$ against logistic-growth with parameters  $\gamma = 2.72\%$, $K = 12,263,457,578 $ Hab. \label{overflow}}
\end{figure}

@@

The resulting graph, while matching the dataset quite nicely, does not make much logical sense. It indicates a very rapid population growth far past 10 billion people before leveling off. This is likely due to the smaller dataset being used - the paper had data going back to 1850 whereas the data used in the simulation only dates back to 1960 due to the lack of accurate population and GDP information dating back much further. However, a big takeaway from this simulation is the similarity of the parameters to the paper's. While the values were clearly different, they were fairly similar and, most notably, $\alpha_3 k_1$ was off by a magnitude of about $10^{-6}$. Based off of this, the original parameters were slightly adjusted manually to see if a more accurate picture could be obtained. The final parameters are $\alpha_1 = 0.3\%$, $\alpha_3 k_1 = \frac{5.2}{10^{6}}\frac{Hab.}{US \$}$, $\alpha_2 = \frac{-55}{10^{17} US\$}$, $\beta_1 = 3.6\%$, and $\beta_2 = \frac{-2}{10^{22} Hab.}$. The resulting graph is remarkably similar to what the report indicates.

@@

The population is predicted to continue rising steadily until it reaches its peak at just under 375 million people in 2060 before decreasing. Unlike the model for the world population though, this decrease does not appear to level off at an equilibrium. This does not make much logical sense since there is no reasonable explanation for the population to steadily decrease that much based on GDP alone. It is likely that the model would perform far better with additional historical data. The paper's simulation of the world population used data reaching back to 1850 whereas the simulation here only used data dating to 1960 due to the lack of credible sources. Being less than a century ago, this is relatively recent and likely deprives the model of crucial information to successfully simulate this system. If additional data were used, the model would likely be similar to what was found with the world population, where an equilibrium value was attained.

\begin{figure}[t]
\includegraphics[width=80mm]{images/us-pop.png}
\includegraphics[width=80mm]{images/us-gdp.png}
\caption{U.S. population and GDP modeled with parameters   $\alpha_1 = 1.07\%$, $\alpha_3 k_1 = \frac{2.76}{10^{7}}\frac{Hab.}{US \$}$, $\alpha_2 = \frac{-1.04}{10^{15} US\$}$, $\beta_1 = 17.9\%$, and $\beta_2 = \frac{-4.63}{10^{10} Hab.}$ against logistic-growth with parameters  $\gamma = 1.71\%$, $K = 632,260,032$ Hab.\label{overflow}}
\end{figure}

@@

The output parameters are $\alpha_1 = 2.22\%$, $\alpha_3 k_1 = \frac{2.77}{10^{6}}\frac{Hab.}{US \$}$, $\alpha_2 = \frac{-8.31}{10^{14} US\$}$, $\beta_1 = 29.0\%$, and $\beta_2 = \frac{-5.21}{10^{9} Hab.}$. The resulting graph for population exhibits a shape similar to that seen with the world simulation, with the population rising, before eventually falling slightly and reaching an equilibrium. GDP is constantly exponentially increasing.

\begin{figure}[t!]
\includegraphics[width=80mm]{images/sa-pop.png}
\includegraphics[width=80mm]{images/sa-gdp.png}
\caption{South Africa population and GDP modeled with parameters $\alpha_1 = 2.22\%$, $\alpha_3 k_1 = \frac{2.77}{10^{6}}\frac{Hab.}{US \$}$, $\alpha_2 = \frac{-8.31}{10^{14} US\$}$, $\beta_1 = 29.0\%$, and $\beta_2 = \frac{-5.21}{10^{9} Hab.}$ against logistic-growth with parameters   $\gamma = 3.85\%$, $K = 79,007,800$ Hab. \label{overflow}}
\end{figure}

>>>>

\begin{figure}[t!]
\includegraphics[width=80mm]{images/world-pop.png}
\includegraphics[width=80mm]{images/world-gdp.png}
\caption{World population and GDP modeled with parameters  $\alpha_1 = 0.3\%$, $\alpha_3 k_1 = \frac{5.2}{10^{6}}\frac{Hab.}{US \$}$, $\alpha_2 = \frac{-55}{10^{17} US\$}$, $\beta_1 = 3.6\%$, and $\beta_2 = \frac{-2}{10^{22} Hab.}$ against logistic-growth with parameters  $\gamma = 2.72\%$, $K = 12,263,457,578 $ Hab. }
\label{fig:world}
\end{figure}

@@

The resulting graph as seen in figure \ref{fig:world}, while matching the dataset quite nicely, does not make much logical sense. It indicates a very rapid population growth far past 10 billion people before leveling off. This is likely due to the smaller dataset being used - the paper had data going back to 1850 whereas the data used in the simulation only dates back to 1960 due to the lack of accurate population and GDP information dating back much further. However, a big takeaway from this simulation is the similarity of the parameters to the paper's. While the values were clearly different, they were fairly similar and, most notably, $\alpha_3 k_1$ was off by a magnitude of about $10^{-6}$. Based off of this, the original parameters were slightly adjusted manually to see if a more accurate picture could be obtained. The final parameters are $\alpha_1 = 0.3\%$, $\alpha_3 k_1 = \frac{5.2}{10^{6}}\frac{Hab.}{US \$}$, $\alpha_2 = \frac{-55}{10^{17} US\$}$, $\beta_1 = 3.6\%$, and $\beta_2 = \frac{-2}{10^{22} Hab.}$. The resulting graph is remarkably similar to what the report indicates.

@@

\begin{figure}[t]
\includegraphics[width=80mm]{images/us-pop.png}
\includegraphics[width=80mm]{images/us-gdp.png}
\caption{U.S. population and GDP modeled with parameters   $\alpha_1 = 1.07\%$, $\alpha_3 k_1 = \frac{2.76}{10^{7}}\frac{Hab.}{US \$}$, $\alpha_2 = \frac{-1.04}{10^{15} US\$}$, $\beta_1 = 17.9\%$, and $\beta_2 = \frac{-4.63}{10^{10} Hab.}$ against logistic-growth with parameters  $\gamma = 1.71\%$, $K = 632,260,032$ Hab.}
\label{fig:us}
\end{figure}

The resulting graph can be seen in figure \ref{fig:us}. The population is predicted to continue rising steadily until it reaches its peak at just under 375 million people in 2060 before decreasing. Unlike the model for the world population though, this decrease does not appear to level off at an equilibrium. This does not make much logical sense since there is no reasonable explanation for the population to steadily decrease that much based on GDP alone. It is likely that the model would perform far better with additional historical data. The paper's simulation of the world population used data reaching back to 1850 whereas the simulation here only used data dating to 1960 due to the lack of credible sources. Being less than a century ago, this is relatively recent and likely deprives the model of crucial information to successfully simulate this system. If additional data were used, the model would likely be similar to what was found with the world population, where an equilibrium value was attained.

@@

\begin{figure}[t!]
\includegraphics[width=80mm]{images/sa-pop.png}
\includegraphics[width=80mm]{images/sa-gdp.png}
\caption{South Africa population and GDP modeled with parameters $\alpha_1 = 2.22\%$, $\alpha_3 k_1 = \frac{2.77}{10^{6}}\frac{Hab.}{US \$}$, $\alpha_2 = \frac{-8.31}{10^{14} US\$}$, $\beta_1 = 29.0\%$, and $\beta_2 = \frac{-5.21}{10^{9} Hab.}$ against logistic-growth with parameters   $\gamma = 3.85\%$, $K = 79,007,800$ Hab.}
\label{fig:sa}
\end{figure}

The output parameters are $\alpha_1 = 2.22\%$, $\alpha_3 k_1 = \frac{2.77}{10^{6}}\frac{Hab.}{US \$}$, $\alpha_2 = \frac{-8.31}{10^{14} US\$}$, $\beta_1 = 29.0\%$, and $\beta_2 = \frac{-5.21}{10^{9} Hab.}$. The resulting graph, as seen in figure \ref{fig:sa} population exhibits a shape similar to that seen with the world simulation, with the population rising, before eventually falling slightly and reaching an equilibrium. GDP is constantly exponentially increasing.
